\chapter{SYSTEM TESTING}

\section{Introduction}
Testing is a critical phase in the development life cycle that ensures the system meets the specified requirements and is free of errors. For the Staff Presence Verification system, testing was categorized into unit, integration, and system-level testing.

\section{Test Plan}
The test plan defines the scope and approach for verifying the BLE detection logic and the web dashboard stability.
\begin{enumerate}
    \item \textbf{Unit Testing}: Testing individual modules like RSSI filtering and database connection.
    \item \textbf{Integration Testing}: Ensuring ESP32 data correctly triggers updates in React components.
    \item \textbf{Acceptance Testing}: Validating that the administrator can generate accurate reports.
\end{enumerate}

\section{Test Case Scenarios}
The following table summarizes the key test cases performed during the system verification.

\begin{table}[H]
\centering
\caption{System Test Case Document}
\begin{tabular}{|c|p{4cm}|p{4cm}|p{3cm}|c|}
\hline
\textbf{S.No} & \textbf{Test Description} & \textbf{Expected Result} & \textbf{Actual Result} & \textbf{Status} \\ \hline
1 & Staff enters room with BLE Tag & Detection array matches UUID after 5s scan & Payload sent to Node.js & Pass \cr \hline
2 & Staff leaves room & RSSI drops below -95dBm & Tag ignored; Status stale & Pass \cr \hline
3 & Attendance Timeout Check & 5min without signal triggers checkout & Status updates to 'Left' & Pass \cr \hline
4 & Mobile GATT Verification & Mobile App writes UUID to BLE Server & Verified as 'mobile_verification' & Pass \cr \hline
5 & WiFi disconnection & ESP32 \texttt{WL\_CONNECTED} loop blocks & Auto-Reconnects when available & Pass \cr \hline
\end{tabular}
\end{table}

\section{Performance Testing}
Performance testing revealed that the Express.js backend efficiently processes the JSON array of detected tags using the MongoDB \texttt{findOne()} aggregation, returning a 200 OK status without bottlenecking the ESP32.
